---
title: "Bayesian Statistics- HW4"
author: "魏上傑"
date: "2023-06-17"
output:
  pdf_document:
    latex_engine: xelatex
  html_document:
    df_print: paged
mainfont: Times New Roman
fontsize: 12pt
papersize: a4
geometry: margin=1.5cm
lang: "zh-tw"
documentclass: ctexart
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(ggplot2)
require(gridExtra)
require(ProbBayes)
require(tidyverse)
crcblue <- "#2905a1"
```

1. 

(a).

```{r}
par(mfrow=c(1, 2))
par(mai=c(0.9,0.9,0.1,0.1))

# Chrystal's prior
X <- rgamma(n = 100, shape = 70, scale = 1/10)
hist(X, prob=TRUE, breaks=10, main=NULL)
x <- seq(min(X), max(X), length.out=100)
lines(x, dgamma(x, shape=70, scale=1/10), col="red")

# Danny's prior
Y <- rgamma(n=100, shape= 33.3, scale = 1/3.3)
hist(Y, prob=TRUE, breaks=10, main=NULL)
y <- seq(min(Y), max(Y), length.out=100)
lines(y, dgamma(y, shape=33.3, scale=1/3.3), col="red")
```

(b).

```{r}
cat("Chrystal's belief on the average ER visits is",mean(X),"\n")
cat("Danny's belief on the average ER visits is", mean(Y),"\n")
var(X)
var(Y)
cat("Chrystal is more confident of her best guess at the average number of ER visits")
```
(c).

```{r}
# Chrystal's 90% credible interval
qgamma(c(0.05, 0.95), shape = 70, scale = 1/10)
# Danny's 90% credible interval
qgamma(c(0.05, 0.95), shape= 33.3, scale=1/3.3)
```

(d).

Need to increase the variance.

2. 

(a). 

```{r}
n <- 7
y.sum <- 8+6+6+9+8+9+7

#Chrystal's 95% posterior credible interval
qgamma(c(0.025, 0.975), shape = 70+y.sum, scale = 1/(10+n))

#Danny's 95% posterior credible interval
qgamma(c(0.025, 0.975), shape = 33.3+y.sum, scale = 1/(3.3+n))


```

(b).

To assess if the statement "the average number of ER visits during any evening hour does not exceed 6" is reasonable, one simply computes its posterior probability, $\Pr (\lambda\ge 6|\alpha_n, \beta_n)$

```{r}
1-pgamma(6, shape = 70+y.sum, scale = 1/(10+n))
1-pgamma(6, shape = 33.3+y.sum, scale = 1/(3.3+n))
```

The probability is very high, so one would conclude that this statement is unlikely to be true.

(c).

```{r}
# use Chrystal's posterior to predict
S <- 7
pred_mu_sim <- rgamma(S, shape = 70+y.sum, rate = 10+n) # sample mu from posterior
pred_y_sim <- rpois(S, lambda = pred_mu_sim)
pred_y_sim
```

```{r}
# use Danny's posterior to predict
pred_mu_sim <- rgamma(S, shape = 33.3+y.sum, rate = 3.3+n)
pred_y_sim <- rpois(S, lambda = pred_mu_sim)
pred_y_sim
```

3. 

Theoretical exercises. Do on your own.


4.

Interpret on your own.